%-------------------------
% Research CV – Muhammad Hamza
% Clean, professional, ATS-parsable
%-------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{fontawesome5}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage[usenames,dvipsnames]{color}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\input{glyphtounicode}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Margins
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Section formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

\pdfgentounicode=1

%-------------------------
% Custom commands

\newcommand{\resumeItem}[1]{
  \item\small{#1 \vspace{-2pt}}
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small #3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeSubSubheading}[2]{
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small #1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeProjectHeading}[2]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \small #1 & #2 \\
    \end{tabular*}\vspace{-5pt}
}

\renewcommand{\labelitemii}{$\circ$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
\begin{document}

%----------HEADING-----------------
\noindent
\begin{tabularx}{\textwidth}{@{} X r @{}}
  \textbf{\Large Muhammad Hamza} &
  \href{mailto:muhammadhamza@kgpian.iitkgp.ac.in}{\faEnvelope\;muhammadhamza@kgpian.iitkgp.ac.in} \\[2pt]
  \href{https://hamza-faarooq.github.io/}{\small hamza-faarooq.github.io} &
  \href{https://github.com/hamza-faarooq}{\faGithub\;hamza-faarooq} \quad
  \href{https://www.linkedin.com/in/muhammad-hamza-iftekhar-hamid-784903204/}{\faLinkedin\;LinkedIn} \\
\end{tabularx}

%-----------EDUCATION-----------------
\section{Education}
\resumeSubHeadingListStart

\resumeSubheading
{Indian Institute of Technology, Kharagpur}{Kharagpur, India}
{Dual Degree (B.Tech \& M.Tech) in Chemical Engineering \;|\; Micro in \textbf{AI \& Applications}}{Aug 2023 -- May 2028}
\resumeItemListStart
\resumeItem {CGPA: \textbf{8.68/10.00}}
\resumeItemListEnd

\resumeSubHeadingListEnd

%-----------RESEARCH EXPERIENCE-----------------
\section{Research Experience}
\resumeSubHeadingListStart

  \resumeSubheading
    {University of Manchester  \normalfont{(Supervisor: Prof.\ Anirbit Mukherjee)}}{Manchester, UK}
    {Research Intern – Scientific Machine Learning}{Oct 2025 -- Present}
  \resumeItemListStart
\resumeItem{Designed and implemented \textbf{DeepONet-based neural operators on spherical domains}, modeling heat PDE dynamics with \textbf{$L_{max}=100+$ harmonic modes}, achieving stable spectral evolution and accurate eigenvalue decay.}
\resumeItem{Engineered \textbf{geometry-aware architectures} combining \textbf{Graph CNN (HEALPix, $\sim$196K nodes)} \& \textbf{SIREN trunk networks}, mitigating spectral bias and enabling high-frequency function reconstruction on $S^2$ manifolds.}

\resumeItem{Developed \textbf{spectral-domain operator learning pipelines} using \textbf{spherical harmonic transforms (SHT)}, ensuring alias-free representations up to \textbf{$L_{max}=511$} and preserving orthogonality for robust PDE generalization.}

\resumeItem{Conducted \textbf{a priori vs a posteriori validation} of learned operators, demonstrating physically-consistent models improve long-time stability by \textbf{reducing spectral drift \& error accumulation} in iterative time evolution.}

\resumeItem{Optimized \textbf{scalable training pipelines} via \textbf{stochastic spatial sampling (5K/196K nodes per step)}, significantly reducing the computational cost by \textbf{$\sim$40×} while preserving convergence and operator fidelity.}

\resumeItem{Investigated \textbf{spectral operator discovery parallels} between turbulent \textbf{Navier–Stokes closure models and spherical PDEs}, analyzing higher-order differential learning for stable and robust neural operator generalization}  \resumeItemListEnd

  \resumeSubheading
    {Indian Institute of Management, Ranchi \normalfont{(Supervisor: Prof.\ Sobhan Sarkar)}}{Remote}
    {Research Intern – Multimodal Machine Learning}{April 2025 -- Aug 2025}
  \resumeItemListStart
\resumeItem{Developed a scalable, robust, end-to-end \textbf{multimodal learning pipeline} integrating textual and visual data, efficiently processing \textbf{54,000+ IMDB reviews} and \textbf{30,000+ movie frames} across \textbf{10+ films} from 4+ genres.}

\resumeItem{Designed and benchmarked architectures combining \textbf{BERT embeddings}, \textbf{Vision Transformer (ViT)} features, and \textbf{CLIP-based alignment} under multiple fusion strategies for improved multimodal performance evaluation.}  

\resumeItem{Evaluated the models on \textbf{five rigorous performance metrics} (Accuracy, F1, AUC, MAP, NDCG), achieving exceptionally strong peak performance of \textbf{0.875 Accuracy} and \textbf{0.982 NDCG}, best on \textbf{'Interstellar'} dataset.}
\resumeItem{Introduced the \textbf{Dempster--Shafer evidential fusion} to explicitly model cross-modal uncertainty \& belief conflict, enabling adaptive, confidence-weighted multimodal predictions improving robustness over baseline fusion methods.}
\resumeItem{Conducted a \textbf{comprehensive comparative analysis of multimodal fusion strategies}, evaluating the \textbf{concatenation, attention-based, and evidential approaches} to model cross-modal interactions, significantly improving ranking performance, semantic alignment, and \textbf{uncertainty-aware decision-making robustness}}
\resumeItemListEnd

  \resumeSubheading
    {University of Oulu, Finland  \normalfont{(Supervisor: Prof.\ Shuai Lee)}}{Remote}
    {Research Intern – Constrained Quadratic Optimisation}{May 2025 -- June 2025}
  \resumeItemListStart
\resumeItem{Formulated \& solved constrained quadratic programs for \textbf{pruning Spiking Neural Networks (SNNs),} achieving \textbf{87.5\% spatial sparsity} while preserving accuracy, \textbf{using CVXPY} with biologically inspired constraints.}
\resumeItem{Integrated criticality-based neuron selection into pruning pipeline using layer-wise importance scores, \textbf{retaining top 112 out of 784 neurons}, and visualizing accuracy versus criticality thresholds to guide structured pruning.}
    \resumeItem{\textbf{Implemented KL-divergence-based temporal sparsity regularization}, penalizing redundant spiking across timesteps and integrating it as a weighted constraint in CVXPY to jointly reduce spatial and temporal redundancy.}
    \resumeItem{\textbf{Conducted experiments on MNIST and CIFAR-10 with rate encoding}, benchmarking the \textbf{accuracy degradation under varying sparsity levels} and \textbf{comparing  the SNN pruning to the ANN baselines.}}
  \resumeItemListEnd

\resumeSubHeadingListEnd

%-----------PUBLICATIONS-----------
\section{Publications \& Patents}
\vspace{4pt}

\noindent\begin{tabularx}{\textwidth}{@{} l X @{}}
  \small\textit{Legend:} & \small C\,=\,Conference, \; J\,=\,Journal, \; P\,=\,Patent, \; A\,=\,Accepted \\[6pt]

  \textbf{[C.1, A]} &
  \textbf{Muhammad Hamza et al.} (2025). Movie Review Helpfulness Prediction using Semantic Alignment \& Evidential Fusion-based Multimodal Approach. \textit{International Joint Symposium on Artificial Intelligence and Natural Language Processing (iSAI-NLP) 2025}; also accepted at \textit{ICEEE--IEEE Conference 2026}. \\
\end{tabularx}

%-----------PROJECTS-----------------
\section{Projects}
\resumeSubHeadingListStart

  \resumeProjectHeading
    {\textbf{Physics-Informed Generative CFD Solver} \textbar\ \emph{PyTorch, Neural Operators, DDPM}}{2026}
  \resumeItemListStart

    \resumeItem{Engineered a \textbf{physics-informed generative pipeline} (PIKAN-FNO) integrating Kolmogorov--Arnold Networks with Fourier Neural Operators to learn solution operators for steady-state Navier-Stokes PDEs with high fidelity.}
    \resumeItem{Implemented a \textbf{Denoising Diffusion Probabilistic Model (DDPM)} with custom U-Net backbone to generate continuous fluid flow fields, enabling data-driven initialization for downstream \textbf{physics-constrained refinement.}}
    \resumeItem{Resolved \textbf{PINN training instability} by designing an Ansatz for exact boundary condition enforcement and applying \textbf{adaptive loss balancing between PDE residuals and constraint terms} for stable convergence.}
    \resumeItem{Conducted \textbf{rigorous model evaluation} using L2 error metrics and qualitative flow-field analysis, demonstrating strong physical consistency and generalization across unseen boundary conditions.}

  \resumeItemListEnd



  \resumeProjectHeading
    {\textbf{Robustness and Calibration Analysis of Transformer Classifiers } $\mid$ \emph{PyTorch, Transformers}}{2025}
  \resumeItemListStart
    \resumeItem{Fine-tuned \textbf{DistilBERT on 67k-sample SST-2 dataset}, achieving \textbf{90\% validation accuracy} \& evaluated robustness under token perturbations, \textbf{observing 38\% relative ECE increase} despite accuracy degradation.}
    
\resumeItem{Performed reliability analysis using \textbf{ECE (0.074 → 0.103)} and reliability diagrams, revealing a \textbf{systematic overconfidence \& confidence–accuracy misalignment} of softmax outputs under simulated distribution shifts.} 
  \resumeItemListEnd


\resumeProjectHeading
{\textbf{{Handwritten Digit Recognition with Neural Networks from scratch}
} $\mid$ \emph{Numpy}}{2024}\resumeItemListStart
\resumeItem{Achieved a \textbf{test accuracy of 91.8\% and a loss of 0.007} in handwritten digit classification using a extensible and scalable feedforward neural network architecture designed \& \textbf{implemented completely from scratch.}}
\resumeItem{Engineered and optimized training pipeline, consistently reducing validation loss \& improving prediction accuracy.}

\resumeItemListEnd

\resumeSubHeadingListEnd


%-----------SCHOLASTIC ACHIEVEMENTS-----------
\section{Scholastic Achievements}
\resumeSubHeadingListStart
  \resumeItemListStart
    \resumeItem{Ranked in the \textbf{top 0.26\%} of \textbf{1.2M+ candidates} in the \textbf{Joint Entrance Examination (JEE) Main, 2023.}}
    \resumeItem{Ranked in the \textbf{top 3\%} among \textbf{200K+ candidates} in \textbf{Joint Entrance Examination (JEE) Advanced, 2023.}}
  \resumeItemListEnd
\resumeSubHeadingListEnd

%-----------COMPETITIONS-----------
\section{Competitions}
\resumeSubHeadingListStart

  \resumeSubheading
    {Optiver Trade-athon 2026}{\textbf{Optiver}}{March 2026}{}
  \resumeItemListStart
    \resumeItem{Secured \textbf{Rank 23} amongst IIT competitors in \textbf{Optiver's annual Trade-athon}, a competitive quantitative trading challenge involving real-time market-making and pricing strategies under dynamic simulated market conditions.}
  \resumeItemListEnd

  \resumeSubheading
    {Convolve 4.0 -- Pan-IIT AI/ML Hackathon}{\textbf{Quadeye Securities/IIT Madras}}{Feb 2026}{}
  \resumeItemListStart
\resumeItem{Secured \textbf{All India Rank 102} in the highly competitive screening ML/AI assessment amongst \textbf{3,000+ students.}}
\resumeItemListEnd

  \resumeSubheading
    {NK Securities Research Hackathon 2025}{\textbf{NK Securities Research}}{Jun 2025}{}
  \resumeItemListStart
    \resumeItem{Secured \textbf{Global Rank 198/6000+} and achieved an ultra-low MSE of \textbf{9.8e-5} in implied volatility prediction using advanced feature engineering, \textbf{outperforming standard ML baselines by 10$\times$ in accuracy.}}
    \resumeItem{Developed a high-frequency volatility forecasting pipeline using \textbf{178K+ records} and \textbf{90+ engineered features}.}
  \resumeItemListEnd

  \resumeSubheading
    {Convolve 3.0 -- Pan-IIT AI/ML Hackathon}{\textbf{IIT Guwahati}}{Jan 2025}{}
  \resumeItemListStart
    \resumeItem{Secured \textbf{Top 50 among 4,000+ teams} by developing a data-driven framework for loan default prediction under real-world constraints, demonstrating robust generalization and decision-aware modeling.}
    \resumeItem{Designed an \textbf{XGBoost-based predictive model} with robust preprocessing for missing data and severe class imbalance, achieving \textbf{0.85 precision} and \textbf{0.91 F1-score} for reliable risk-sensitive predictions.}
  \resumeItemListEnd

\resumeSubHeadingListEnd

%-----------TECHNICAL SKILLS-----------
\section{Technical Skills}
\resumeSubHeadingListStart
  \item\small{
    \textbf{Languages \& Frameworks:} Python, C++, C, SQL, PyTorch, TensorFlow, NumPy, Pandas, Scikit-learn, SciPy \\
    \textbf{ML / AI:} Neural Operators (FNO, DeepONet), PINNs, Diffusion Models, Transformers (ViT, BERT, CLIP), SNNs \\
    \textbf{Tools:} Git, \LaTeX
  }
\resumeSubHeadingListEnd


% \section{\textbf{Coursework Information}}

% \begin{itemize}[leftmargin=0.05in, label={}]
% \item \textbf{Mathematics and Computer Science:} Artificial Intelligence: Foundations and Applications, Programming and Data Structures (Theory and Laboratory), Linear Algebra, Advanced Calculus, Transform Calculus, Numerical and Complex Analysis, Computer-Aided Process Engineering
% \vspace{-5pt}
% \item \textbf{MOOCs:} Classical Machine Learning, Neural Networks and Deep Learning, Improving Deep Neural Networks, Convolutional Neural Networks, Natural Language Processing, Large Language Models
% \end{itemize}

%-------------------------------------------
\end{document}
